An efficient acyclic contact planner for multiped robots

### 2016-02-03 body/abstract.tex
final submission

We address the issue of automatically planning discrete sequences of contacts
that characterize a multiped motion in a constraining environment.
To overcome the inherent combinatorial difficulty of this problem, we separate it
in two sub-problems: first, planning a guide path for the root of the robot and
then, generating contacts along this path.
We propose theoretical contributions for each sub-problem, and implement them
as a complete contact planner. 
First, we generate a guide root path without explicitly computing configurations in static equilibrium, as opposed to previous approaches.
To do this we propose to approximate the space of equilibrium configurations using the ``reachability condition".
Verifying this condition boils down to performing simple collision tests and, in the considered environments, this approximation turns out to be accurate.
Then, we propose an algorithm to extend a root guide path into a sequence of equilibrium configurations.
A sampling-based approach is used to generate candidate contacts.
The best candidate is chosen according to user-defined heuristics. We propose four such heuristics, used in all our experiments, and 
designed for generic robotic applications.
These contributions are implemented within the open-source software Hpp. Our framework allows us to generate complex multi-contact sequences in at most a few seconds for a large variety of robots, including HRP-2, HyQ, and a dexterous hand.

>>>>

We present a framework capable of producing contact plans describing complex multiped motions (including humanoid): standing up, climbing stairs using a handrail, crossing rubble and getting out of a car. Our framework answers a need demonstrated at the Darpa Robotics Challenge,
where the lack of an automatic acyclic contact planner was recognized a major issue.

Our novel key idea is the reachability condition. Informally, it verifies that the root configuration of a robot is ``close, but not too close'' from obstacles: close to allow contact creation, not too close to avoid collision. With this approximation of the space of admissible root configurations we decompose the hard contact planning problem into simpler sub-problems: first, to plan a guide path for the root without considering the whole-body configuration; then, to generate a discrete sequence of whole-body configurations in static equilibrium along this path. The reachability condition turns the high-dimensional computation of the guide into a collision checking problem, solved in less than a few seconds. Then a deterministic contact selection algorithm tackles the combinatorial issue of  generating of a discrete sequence along the guide path. Several innovations make it computationally efficient: a criterion for verifying static equilibrium, and a set of heuristics used to enforce desirable properties on the configuration.

Our approach results from the pragmatic choice of favoring efficiency over exhaustiveness, justified empirically: in a few seconds, with satisfying success rates, we generate complex contact plans for various scenarios and robots, namely HRP-2, HyQ, and a dexterous hand.

### 2016-02-03 body/results.tex
final submission

\begin{table*}
\centering
\begin{tabular}{ c | c | c | c | c }
 Scenario & Method  & Guide Path & Contact sequence & Interpolation \\
 \hline
   \multirow{3}{*}{Stair 20 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}5.42 min}  \\
							 & \textbf{Ours} + \citeauthor{Carpentier2016}  & $\mathbf{<}$ \textbf{2 ms} & $\mathbf{<}$ \textbf{14 ms} & $ <$ 2s \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{3}{*}{Stair 30 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}4.08 min}  \\
							 & \textbf{Ours}  & $\mathbf{<}$ \textbf{2 ms} & $\mathbf{<}$ \textbf{14 ms} & X \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{3}{*}{Stair 40 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}10.08 min}  \\
							 & \textbf{Ours}  & \textbf{5 s} & \textbf{3 s} & X \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{2}{*}{Table (car) egress} & \citeauthor{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 10 min & \multicolumn{2}{c}{\cellcolor{blue!25}3.5 hours}  \\
							 & \textbf{Ours}  & \textbf{529 ms} & \textbf{$<$ 22 s} & X \\
 \end{tabular}
\caption{Comparison between the computation times obtained by our method and previous ones, for each phase of the planning. Blue cells indicate merged times (because no
precise time information is available). The X symbol indicates scenarios where we did not address the interpolation of the contact sequence.}
\label{tab:compprev}
\quad
\end{table*}
\subsection{Comparison with previous work} \label{sec:compa}
Comparing our method with others is not trivial. Indeed, we focus on the contact planning aspect of the problem, and do not address
the interpolation in this paper, nor provide a strong guarantee that our sequences can be interpolated.
Most papers in the literature only provide the overall computation
time, without specifying how much is spent in each phase. Moreover, there exists no off-the-shelf implementation
of the related methods. Due to their complexity, we cannot afford the development resources required to reimplement them.

>>>>

\begin{table*}
\centering
\begin{tabular}{ c | c | c | c | c }
 Scenario & Method  & Guide Path & Contact sequence & Interpolation \\
 \hline
   \multirow{3}{*}{Stair 20 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}5.42 min}  \\
							 & \textbf{Ours} + \citeauthor{Carpentier2016}  & \textbf{5 ms} & $\mathbf{<}$ \textbf{18 ms} & $ <$ 2s \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{3}{*}{Stair 30 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}4.08 min}  \\
							 & \textbf{Ours}  & \textbf{5 ms} & $\mathbf{<}$ \textbf{18 ms} & X \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{3}{*}{Stair 40 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{\cellcolor{blue!25}10.08 min}  \\
							 & \textbf{Ours}  & \textbf{5 s} & \textbf{3 s} & X \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{\cellcolor{blue!25}2 to 10 min} \\
 \hline
   \multirow{2}{*}{Table (car) egress} & \citeauthor{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 10 min & \multicolumn{2}{c}{\cellcolor{blue!25}3.5 hours}  \\
							 & \textbf{Ours}  & \textbf{529 ms} & \textbf{$<$ 22 s} & X \\
 \end{tabular}
\caption{Comparison between the computation times obtained by our method and previous ones, for each phase of the planning. Blue cells indicate merged times (because no
precise time information is available). The X symbol indicates scenarios where we did not address the interpolation of the contact sequence.}
\label{tab:compprev}
\quad
\end{table*}
\subsection{Comparison with previous work} \label{sec:compa}
Comparing our method with others is not trivial. Indeed, we focus on the contact planning aspect of the problem, and do not address
the interpolation in this paper, nor provide a strong guarantee that our sequences can be interpolated.
Most papers in the literature only provide the overall computation
time, without specifying how much is spent in each phase. Moreover, there exists no off-the-shelf implementation
of the related methods. Due to their complexity, we cannot afford the development resources required to reimplement them.

### 2016-02-03 main.tex
nicolas comments
final version !
final submission

\setcounter{secnumdepth}{3}
\usepackage{moreverb,url}
\usepackage{makeidx}         % allows index generation
\usepackage[percent]{overpic}
\usepackage{array} 
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{color}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{multirow}

@@

\section{Introduction}
\input{body/intro}
\input{body/overview}
\input{body/rbprm}
\input{body/contact}
\input{body/equilibrium}
\input{body/efort}
\input{body/results}
\input{body/conclusion}
\section*{Acknowledgements}
This research is supported by Euroc (project under FP7 Grant Agreement  608849),  Entracte (ANR  grant  agreement  13-CORD-002-01),
the ARO Contract W911NF-14-1-0437 and the NSF award 1305286. %, and a grant from Boeing.
\theendnotes

>>>>

\setcounter{secnumdepth}{3}
\usepackage{moreverb,url}
\usepackage{makeidx}         % allows index generation
\usepackage[percent]{overpic}
\usepackage{array} 
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{color}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage[table]{xcolor}

@@

\section{Introduction}
\input{body/intro}
\input{body/overview}
\input{body/rbprm}
\input{body/contact}
\input{body/equilibrium}
\input{body/efort}
\input{body/results}
\input{body/conclusion}
\section*{Acknowledgements}
This research is supported by Euroc (project under FP7 Grant Agreement  608849);  Entracte (ANR  grant  agreement  13-CORD-002-01);
the ARO Contract W911NF-14-1-0437; and the NSF award 1305286. %, and a grant from Boeing.
%~ \theendnotes
